% !TeX spellcheck = en_US
%%
%% Text of diploma thesis
%%
%% Tomáš Zítka
%%
\section{Example PDEs}
In the following examples we will busy ourselves solving following equations:\\
Transient advection equation in one resp. two dimensions
\begin{equation}
    \label{eq:ex_advection}
    \pdiff{p}{t} + a\pdiff{p}{x} = 0,
\end{equation}
resp.
\begin{equation}
    \pdiff{p}{t} + \vec{a}\cdot\nabla{p} = 0.
\end{equation}
After applying discretizations devised in \Cref{ch:theory} we obtain both 
equations in the same form
\begin{equation}
         \fdiff{P^k_i}{t}(t)\elint\psi_i\psi_j 
         - \int_{T^k} \vec{a}P_i^k\psi_i\cdot\nabla\psi_j 
         + \int_{\partial{T^k}} \vec{n}
        \cdot f^{*} (p_{in}, p_{out})\cdot\psi_j = 0
\end{equation}
Equation expressed in \sfepy{} declarative notation can be found in Listing 
\ref{lst:advection} below;
\setcounter{lstannotation}{0}
\begin{lstlisting}[language=Python, caption={Advection equation}
\label{lst:advection}]
equations = {'Advection': 
  # transient
  "dw_volume_dot.i.Omega(v, p)"
 
  "- dw_s_dot_mgrad_s.i.Omega(a.val, p[-1], v)" /*!\lann{lsta:history}!*/
  "+ dw_dg_advect_laxfrie_flux.i.Omega(a.flux, a.val, v, p[-1])"  /*!\lann{lsta:flux}!*/
  "= 0"
  }
\end{lstlisting}
\begin{itemize}
    \item[\ref{lsta:history}] \pysauce{"p[-1]"} ensures the variable object 
    stores history one step backwards in time facilitating forward nature of 
    time stepping solvers,
    \item[\ref{lsta:flux}] \pysauce{"a.flux"} is an optional material argument 
    representing $\alpha$ coefficient in \eqref{eq:lax-frieflux}.
\end{itemize}
Laplace equation in two dimensions
\begin{equation}
    \label{eq:ex_laplace}
    - D \left(\pdiff{^2 p}{x^2} + \pdiff{^2 p}{y^2} \right)= 0.
\end{equation}
Employing symmetric discretization of diffusion term and adding diffusion 
penalty term yields equation in the form
\begin{multline}
    \elint D\nabla P^k_i\psi_i \nabla \psi_j
    - \elbint D \manglebracs{P^k_i\nabla\psi_i} \cdot \vec{n}\psi_j 
    - \elbint D \frac{\nabla \psi_j}{2}\cdot \vec{n} [P^k_i\psi_i] \\ 
    + \nu \elbint \sigma [P^k_i\psi_i]\psi_j 
    = 0.
\end{multline}
Discretization of this equation using \sfepy{} terms can be found in Listing 
\ref{lst:laplace};\\
Static advection-diffusion equation with right hand side of the form
\begin{equation}
\label{eq:ex_advdiff}
\pdiff{p}{x} + \pdiff{p}{y} - D \cdot \left( \pdiff{^2 p}{x^2} + \pdiff{^2 
p}{y^2} \right) = g,
\end{equation}
i.e.,
\begin{equation}
\vec{a} \cdot \nabla p - D \Delta p = g,
\end{equation}
where $\vec{a} = [1, 1]^T$ is advection velocity, $D$ is diffusion coefficient 
and $g$ is a source function. Combining discretization of the two previous 
equations we obtain discretized form
\begin{multline}
- \int_{T^k} \vec{a}P_i^k\psi_i\cdot\nabla\psi_j 
+ \int_{\partial{T^k}} \vec{n}\cdot f^{*} (p_{in}, p_{out})\cdot\psi_j\\
+ \elint D\nabla P^k_i\psi_i \nabla \psi_j
- \elbint D \manglebracs{P^k_i\nabla\psi_i} \cdot \vec{n}\psi_j
- \elbint D \frac{\nabla \psi_j}{2}\cdot \vec{n} [P^k_i\psi_i] \\
+ \nu \elbint \sigma [P^k_i\psi_i]\psi_j
- \elint g\cdot \psi_j
= 0.
\end{multline}
In \sfepy{} declarative notation this equation has the form presented in 
Listing \ref{lst:advdiff}.
\setcounter{lstannotation}{0}
\begin{lstlisting}[language=Python, caption=Static advection-diffusion equation
\label{lst:advdiff}]
equations = {'adv_diff' :
    # advection
    "- dw_s_dot_mgrad_s.i.Omega(a.val, p, v)"
    "+ dw_dg_advect_laxfrie_flux.i.Omega(a.flux, a.val, v, p)"
    # diffusion
    "+ dw_laplace.i.Omega(D.val, v, p) "
    "- dw_dg_diffusion_flux.i.Omega(D.val, p, v)"
    "- dw_dg_diffusion_flux.i.Omega(D.val, v, p)"
    # penalty
    "+ dw_dg_interior_penalty.i.Omega(D.val, D.cw, v, p)"
    # source
    "- dw_volume_lvf.i.Omega(g.val, v)"
    "= 0"
    }
\end{lstlisting}
Transient viscous Burgers equation in one resp. two dimensions
\begin{equation}
\pdiff{p}{t} + \frac{1}{2}\pdiff{p^2}{x} - D \cdot\pdiff{^2 p}{x^2} = g,
\end{equation}
resp.
\begin{equation}
\label{eq:ex_burgers}
    \pdiff{p}{t} + \frac{1}{2}\left(\pdiff{p^2}{x} + \pdiff{p^2}{y}\right)  - 
    D \cdot \left( \pdiff{^2 p}{x^2} + \pdiff{^2 p}{y^2} \right) 
    = g,
\end{equation}
i.e.,
\begin{equation}
    \pdiff{p}{t} + \nabla \cdot \vec{f}(p) - D\Delta p = g.
\end{equation}
with $\vec{f}(p) = \frac{1}{2}[p^2, p^2]^T = \frac{1}{2}\vec{a} p^2$.
Discretizing using all the terms derived before we get the same form for both 1D and 2D
\begin{multline}
    \fdiff{P^k_i}{t}(t)\elint\psi_i\psi_j 
    - \int_{T^k} \vec{f}(P_i^k\psi_i)\cdot\nabla\psi_j 
    + \int_{\partial{T^k}} \vec{n} \cdot f^{*} (p_{in}, p_{out})\cdot\psi_j\\
    + \elint D\nabla P^k_i\psi_i \nabla \psi_j
    - \elbint D \manglebracs{P^k_i\nabla\psi_i} \cdot \vec{n}\psi_j
    - \elbint D \frac{\nabla \psi_j}{2}\cdot \vec{n} [P^k_i\psi_i] \\
    + \nu \elbint \sigma [P^k_i\psi_i]\psi_j
    - \elint g\cdot \psi_j
    = 0.
\end{multline}
In \sfepy{} declarative notation this equation has the form presented in 
Listing \ref{lst:burgers}.
\setcounter{lstannotation}{0}
\begin{lstlisting}[language=Python, caption=Viscous Burgers equation \label{lst:burgers}]
burg_velo = nm.array([1., 1.])

def f(p):
    return .5*burg_velo * p[..., None] ** 2

def f_d(p):
    return burg_velo * p[..., None]

equations = {'burgers':
    # transient
    "dw_volume_dot.i.Omega(v, p)"
    #  non-linear hyperbolic terms
    "- dw_ns_dot_grad_s.i.Omega(f, f_d, p[-1], v)" /*!\lann{lsta:nl1}!*/
    "+ dw_dg_nonlinear_laxfrie_flux.i.Omega(f, f_d, v, p[-1])" /*!\lann{lsta:nl2}!*/
    #  diffusion
    "+ dw_laplace.i.Omega(D.val, v, p[-1])"
    "- dw_dg_diffusion_flux.i.Omega(D.val, p[-1], v)"
    "- dw_dg_diffusion_flux.i.Omega(D.val, v, p[-1])"
    # penalty
    "+ dw_dg_interior_penalty.i.Omega(D.val, D.Cw, v, p[-1])"
    # source
    "- dw_volume_lvf.i.Omega(g.val, v)"
    " = 0"
    }
\end{lstlisting}
\begin{itemize}
    \item[\ref{lsta:nl1}, \ref{lsta:nl2}] Nonlinear terms require as parameters function 
    and its derivative.
\end{itemize}


\section{Examples}

\paragraph{Measuring convergence} We define convergence rate $r$ as is common 
in literature
\begin{equation}
%    num_order = nm.log(row["diff_l2"] / last_err) \
%    / nm.log(row["h"] ** dim / last_h ** dim)
r = \frac{\log\left(\dfrac{\Ltwonorm{p - p_{h_{n}}}}{\Ltwonorm{p - p_{h_{n-1}}}}\right)}
{\log\left(\dfrac{h^d_{n}}{h^d_{n-1}}\right)}.
\end{equation}
Inspired by \cite{Kucera} we present plots depicting average convergence rate 
over several mesh refinements, this might not be ideal measure of the method 
behavior, nevertheless it still provides us with convenient indicator. 
Accompanied with plots of $L^2$ error it allows us to reason about method over 
several varying parameters, notably it reveals relationship between diffusion 
coefficient and penalty term in examples including diffusion terms.

In \Cref{ex:adv1D} we explore behavior of DG FE method for 1D pure advection, 
time dependent problem with and without limiting, examples \ref{ex:laplace}, 
\ref{ex:quart1}, \ref{ex:quart2} and \ref{ex:quart3} importance of diffusion 
penalty terms. All the test problems studied in Chapter \ref{ch:convergence} 
are specified using this declarative approach, codes can be found in (\todo 
Attachment? \url{https://github.com/zitkat/dg_examples})
